---
title: "Pond Volume and Counts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

How much abatement did water bodies receive? How does this correlative with water body volume? 
```{r}
Ponds2022 <- read_csv("chad_abate_2022_10-27-23.csv",locale=locale(encoding="latin1"))

range(Ponds2022$volume) #three ponds listed with no size or volume

abatequantityNoZeros2022 = Ponds2022$abatequantity[Ponds2022$abatequantity>0] 

abatevolumeNoZeros2022 = Ponds2022$volume[Ponds2022$abatequantity>0] 

range(abatequantityNoZeros2022/abatevolumeNoZeros2022) 

hist(abatequantityNoZeros2022/abatevolumeNoZeros2022)

sort(abatequantityNoZeros2022/abatevolumeNoZeros2022, decreasing = TRUE)[1:100] #one water body with ratio of 2005

abatevolumeNoZeros2022[abatequantityNoZeros2022/abatevolumeNoZeros2022>500] #very different than others, potentially typo, should remove from analysis. Wide shallow waterbopdy that received a lot of abate.

hist(abatequantityNoZeros2022[(abatequantityNoZeros2022/abatevolumeNoZeros2022)<50]/abatevolumeNoZeros2022[(abatequantityNoZeros2022/abatevolumeNoZeros2022)<50]) 
```
Finding:Most received relatively small abatement per unit, one outlier water body that should be excluded 

How do water body counts change for all villages (n=700) from 2022 over the course of the year? 
```{r}
Ponds2022 <- Ponds2022 %>% mutate(Date = as.Date(treatmentdate, format = "%m/%d/%y")) 

Ponds2022$Month = months(as.Date(Ponds2022$treatmentdate))

Ponds2022numeric <- Ponds2022 %>%
  mutate(Month = as.numeric(factor(Month, levels = month.name)))

Ponds2022_summary <- Ponds2022numeric %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

#unique(Ponds2022_summary$village) #700 unique villages

ggplot(Ponds2022_summary, aes(x = Month, y = Count)) +                #plot with individual lines for each village
  geom_line(aes(color = village)) + theme(legend.position="none")

#Ponds2022_summary$village = as.factor(Ponds2022_summary$village)

Villages2022 = gamm(round(Count) ~ s(Month) + s(village, bs = "re"),
                         family = quasipoisson(), correlation = corCAR1(form = ~Month | village),
                         data = Ponds2022_summary) 


Villages2022plot =plot_smooth(Villages2022$gam, view="Month",lwd=2,
                                  transform=exp,se=1, shade=T, #ylim = c(0,50),
                                  ylab="Water Body Count",rug=F,
                                  hide.label=T, n.grid = 100)$fv

ggplot() + geom_line(data = SelectResults2022plot,aes(x=Month, y=fit)) +  #fitted line
  geom_ribbon(data = SelectResults2022plot,aes(x=Month, y=fit, ymin = ll,ymax=ul),alpha=0.2) + #recreates plot smooth
  geom_point(data=SelectVillages2022,aes(x=Month,y=Count)) #raw data
```
Summary: The data is from the Carter Center Guinea Worm Eradication Program 2022 Abate Treatment records. The plot is divided up into three main segments: April – July (beginning of wet season, presumably more, small water bodies form), August-October (wet season continues and flood plain fills), November-March (dry season). There is anecdotal evidence that as the flood plain fills water bodies become too large and therefore "untreatable" and previously treated water bodies become unreachable for treatment, both of which may also be reflected in the decline of water bodies seen later in the wet season. I will verify if these months correspond with what is generally considered the beginning and end of the rainy season.  

Questions: How does water body size change over the course of these three main segments? Is the beginning of the wet season when we see an increase in the number of water bodies characterized by more water bodies of small sizes? Is the latter part of the wet season characterized by larger water bodies?  

Approach: Start with villages that are ground-truthed. Identify changes in the size of specific water bodies over time.  
```{r}
#Look at Bem-Bem 2 first

Villages2022 <- Ponds2022 %>% group_by(watername,zone) #filter by individual villages 

BemBem2 <-  filter(Test,village == "bem-bem 2") 

#Bouta Sidiri (inidividual wb in Bem-Bem 2), look at changes in water body size over time

BoutaSidiri = dplyr::filter(BemBem2, grepl("bouta si",watername))
BoutaSidiri = dplyr::filter(BoutaSidiri, !grepl("bouta simpo",watername))

BoutaSidiri <- BoutaSidiri %>% mutate(Date = as.Date(treatmentdate, format = "%m/%d/%y")) 

BoutaSidiri$Month = months(as.Date(BoutaSidiri$treatmentdate))

BoutaSidiri = BoutaSidiri %>% mutate(Month = as.numeric(factor(Month, levels = month.name)))

#manually rename waternames to fix typos (find easier way to do this?)
waternamecorrected = data.frame(c(rep("Bouta Sidiri 3", times = 2),"BoutaSidiri 1", "Bouta Sidiri 2", rep("Bouta Sidiri 3", times = 2), "Bouta Sidiri 4", rep("Bouta Sidiri 3", times = 2), "Bouta Sidiri 1", "Bouta Sidiri 5", "Bouta Sidiri b", "Bouta Sidiri c", "Bouta Sidiri 5", "Bouta Sidiri 3c", "Bouta Sidiri 4"))
BoutaSidiri = cbind(waternamecorrected,BoutaSidiri)
colnames(BoutaSidiri)[1] = "waternamecorrected" 

library(RColorBrewer)
colourCount = length(unique(BoutaSidiri$waternamecorrected))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))
colrs<- getPalette(colourCount)
names(colrs) <- unique(BoutaSidiri$waternamecorrected)
BoutaSidiri$color <- colrs[BoutaSidiri$waternamecorrected]  
BoutaSidiri$color
                
ggplot() + geom_point(data=BoutaSidiri,aes(x=Month,y=volume), color = BoutaSidiri$color, size = 5) 
```

What villages may have more complete data sets? Look at others from Stephanie's ground truthed list.
```{r}
#Village = Kob
kob2022 = filter(Ponds2022numeric,village == "kob") 
MareCarriereBrique = dplyr::filter(kob2022, grepl("mare c",watername))
MareCarriereBrique = dplyr::filter(MareCarriereBrique, !grepl("zara",watername))
MareCarriereBrique$Month = as.factor(MareCarriereBrique$Month)
ggplot() + geom_point(data=MareCarriereBrique,aes(x=Month,y=volume))

#Skim through other villages to see if there are waterbodies with the same or similiar names each month. Try to identify those that span 12 months.

#Guimeze2022 = filter(Ponds2022numeric,village == "guimeze") Don't seem to be many that fit criteria

#KS2022 = filter(Ponds2022numeric,village == "kolemara sara") Don't seem to be many that fit criteria

#Loumia Center

LC2022 = filter(Ponds2022numeric,village == "loumia centre") 

ggplot() + geom_point(data=LC2022,aes(x=Month,y=volume)) #volume peaks in ~July

LC2022_summary <- LC2022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

ggplot() + geom_point(data=LC2022_summary,aes(x=Month,y=Count)) #count peaks between March and June

```
Check dynamics of specific water bodies in loumia that have consistent names, start with loumia 1,2,3.
