---
title: "Select Chadian Villages"
output: html_document
date: "2024-01-26"
---


```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(mgcv)
library(deSolve)
library(tidyverse)
library(stats)
library(itsadug)
library(data.table)
library(car)
library(broom)
library(ggplot2)
library(cowplot)
```

```{r}
setwd("~/Desktop/Rscripts/Data")
Ponds2022 <- read_csv("chad_abate_2022_10-27-23.csv",locale=locale(encoding="latin1"))

Ponds2022 <- Ponds2022 %>% mutate(Date = as.Date(treatmentdate, format = "%m/%d/%y")) 

Ponds2022$Month = months(as.Date(Ponds2022$treatmentdate))

Ponds2022numeric <- Ponds2022 %>%
  mutate(Month = as.numeric(factor(Month, levels = month.name)))
```

Summarised Villages 
```{r}
#Toukra Mousgoum
TM2022 = filter(Ponds2022numeric,village == "toukra mousgoum") 

TM2022_summary <- TM2022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

colnames(TM2022_summary) = c("Month", "Village", "Count")

TM2022_summary[TM2022_summary=="toukra mousgoum"] <- "Toukra Mousgoum"

TM2022_summary$Village = as.factor(TM2022_summary$Village) 

TM2022Year = data.frame(c(rep("2022", times = 10)))

ToukraMousgoum = cbind(TM2022_summary,TM2022Year)

colnames(ToukraMousgoum) = c("Month","Village", "Count", "Year")

#KobÈ
kob2022 = filter(Ponds2022numeric,village == "kob") 

kob2022_summary <- kob2022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

colnames(kob2022_summary) = c("Month", "Village", "Count")

kob2022_summary[kob2022_summary=="kob"] <- "Kob"

kob2022_summary$Village = as.factor(kob2022_summary$Village) 

KobYear2022 = data.frame(c(rep("2022", times = 12)))

Kob = cbind(kob2022_summary,KobYear2022)

colnames(Kob) = c("Month","Village", "Count", "Year")

#Madjira 1
Madjira12022 = filter(Ponds2022numeric,village == "madjira 1") 

Madjira12022_summary <- Madjira12022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

colnames(Madjira12022_summary) = c("Month", "Village", "Count")

Madjira12022_summary[Madjira12022_summary=="madjira 1"] <- "Madjira 1"

Madjira12022_summary$Village = as.factor(Madjira12022_summary$Village) 

Madjira2022Year = data.frame(c(rep("2022", times = 12)))

Madjira1 = cbind(Madjira12022_summary,Madjira2022Year)

colnames(Madjira1) = c("Month","Village", "Count", "Year")

#Diganali 
Diganali2022 = filter(Ponds2022numeric,village == "diganali") 

Diganali2022_summary <- Diganali2022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

colnames(Diganali2022_summary) = c("Month", "Village", "Count")

Diganali2022_summary[Diganali2022_summary=="diganali"] <- "Diganali"

Diganali2022_summary$Village = as.factor(Diganali2022_summary$Village) 

DiganaliYear2022 = data.frame(c(rep("2022", times = 11)))

Diganali = cbind(Diganali2022_summary,DiganaliYear2022)

colnames(Diganali) = c("Month","Village", "Count", "Year")

#Asso2
Asso22022 = filter(Ponds2022numeric,village == "asso 2") 

Asso22022_summary <- Asso22022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

colnames(Asso22022_summary) = c("Month", "Village", "Count")

Asso22022_summary[Asso22022_summary=="asso 2"] <- "Asso 2"

Asso22022_summary$Village = as.factor(Asso22022_summary$Village) 

Asso2Year2022 = data.frame(c(rep("2022", times = 12)))

Asso2 = cbind(Asso22022_summary,Asso2Year2022)

colnames(Asso2) = c("Month","Village", "Count", "Year")

#LoumiaCenter
LC2022 = filter(Ponds2022numeric,village == "loumia centre") 

write.csv(LC2022, "~/Desktop/Rscripts/Data/LoumiaCentre2022.csv", row.names=FALSE)

LC2022_summary <- LC2022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

colnames(LC2022_summary) = c("Month", "Village", "Count")

LC2022_summary[LC2022_summary=="loumia centre"] <- "Loumia Centre"

LC2022_summary$Village = as.factor(LC2022_summary$Village) 

LCYear2022 = data.frame(c(rep("2022", times = 12)))

LoumiaCentre = cbind(LC2022_summary,LCYear2022)

colnames(LoumiaCentre) = c("Month","Village", "Count", "Year")

#Mailao

Mailao2022 = filter(Ponds2022numeric,village == "mailao") 

Mailao2022_summary <- Mailao2022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

colnames(Mailao2022_summary) = c("Month", "Village", "Count")

Mailao2022_summary[Mailao2022_summary=="mailao"] <- "Mailao"

Mailao2022_summary$Village = as.factor(Mailao2022_summary$Village) 

MailaoYear2022 = data.frame(c(rep("2022", times = 12)))

Mailao = cbind(Mailao2022_summary,MailaoYear2022)

colnames(Mailao) = c("Month","Village", "Count", "Year")

#Kolemara Sara
KS2022 = filter(Ponds2022numeric,village == "kolemara sara") 

KS2022_summary <- KS2022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

colnames(KS2022_summary) = c("Month", "Village", "Count")

KS2022_summary[KS2022_summary=="kolemara sara"] <- "Kolemara Sara"

KS2022_summary$Village = as.factor(KS2022_summary$Village) 

KS2022dates = data.frame(c(rep("2022", times = 12)))

KolemaraSara = cbind(KS2022_summary,KS2022dates)

colnames(KolemaraSara) = c("Month","Village", "Count", "Year")

#Guizeme 

Guimeze2022 = filter(Ponds2022numeric,village == "guimeze") 

Guimeze2022_summary <- Guimeze2022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

colnames(Guimeze2022_summary) = c("Month", "Village", "Count")

Guimeze2022_summary[Guimeze2022_summary=="guimeze"] <- "Guimeze"

Guimeze2022_summary$Village = as.factor(Guimeze2022_summary$Village) 

Guimeze2022year = data.frame(c(rep("2022", times = 12)))

Guimeze = cbind(Guimeze2022_summary,Guimeze2022year)

colnames(Guimeze) = c("Month","Village", "Count", "Year")

#Bem Bem-2

bembem2022 = filter(Ponds2022numeric,village == "bem-bem 2") 

bembem2022_summary <- bembem2022 %>%
  group_by(Month, village) %>%
  summarise(Count = n()) 

colnames(bembem2022_summary) = c("Month", "Village", "Count")

bembem2022_summary[bembem2022_summary=="bem-bem 2"] <- "Bem-Bem 2"

bembem2022_summary$Village = as.factor(bembem2022_summary$Village) 

BemBem2022Year = data.frame(c(rep("2022", times = 12)))

BemBem2 = cbind(bembem2022_summary,BemBem2022Year)

colnames(BemBem2) = c("Month","Village", "Count", "Year")

```

Cleaning data so that point of entry/sections are not double counted. 

Kolemara Sara 
```{r}
unique_sections_per_month <- KS2022 %>%
  group_by(Month) %>%
  summarise(unique_sections = unique(sectionname))

entire <- KS2022 %>%
  filter(treatmenttype == 'entire') %>%
  group_by(Month) %>%
  summarise(waternames = unique(watername))

# January - 3
# Feb - 1
# March - 1
# April - 2
#May - 3
#June - 4
#July - 4
#August - 5
#Sept - 2
#Oct - 5
#Nov - 1
#Dec - 1

revisedcounts = c(3,1,1,2,3,4,4,5,2,5,1,1)
RevisedSummaryKS = cbind(KS2022_summary,revisedcounts)
colnames(RevisedSummaryKS) = c("Month","Village", "Count", "RevisedCount")
```

Loumia Centre
```{r}
unique_sections_per_month_LC <- LC2022 %>%
  group_by(Month) %>%
  summarise(unique_sections = unique(sectionname))

entire_LC <- LC2022 %>%
  filter(treatmenttype == 'entire') %>%
  group_by(Month) %>%
  summarise(waternames = unique(watername))

revisedcountsLC = c(13,15,15,16,16,18,15,16,15,12,12,10)

RevisedSummaryLC = cbind(LC2022_summary,revisedcountsLC)
colnames(RevisedSummaryLC) = c("Month","Village", "Count", "RevisedCount")
```

Bem-Bem 2
```{r}
unique_sections_per_month_bembem2 <- bembem2022 %>%
  group_by(Month) %>%
  summarise(unique_sections = unique(sectionname))

entire_bembem2 <- bembem2022 %>%
  filter(treatmenttype == 'entire') %>%
  group_by(Month) %>%
  summarise(waternames = unique(watername))

#counts for entire 
entireWBBB2 = entire_bembem2 %>% group_by(Month) %>%
  summarise(UniqueWaterNames = n_distinct(waternames))

#manually go through sections until come up with better system
#Jan - 1
#Feb - 1
#March - 2
#April - 2
#May - 2
#June - 3
#July - 0
#August - 1
#Sept - 2
#Oct - 1
#Nov - 1
#Dec - 2

sectionsBB2 = data.frame(c(1,1,2,2,2,3,0,1,2,1,1,2)) 

#revisedcountsbembem2 = cbind(entireWBBB2,sectionsBB2) not working cause different # of rows, do by hand atm

revisedcountsbembem2 = c(5,3,6,4,5,6,8,5,4,1,1,3)
RevisedSummaryBB2 = cbind(bembem2022_summary,revisedcountsbembem2)
colnames(RevisedSummaryBB2) = c("Month","Village", "Count", "RevisedCount")
```

Asso 2
```{r}
unique_sections_per_month_Asso2 <- Asso22022 %>%
  group_by(Month) %>%
  summarise(unique_sections = unique(sectionname))

entire_Asso2 <- Asso22022 %>%
  filter(treatmenttype == 'entire') %>%
  group_by(Month) %>%
  summarise(waternames = unique(watername))

#counts for entire 
entireAssoo2counts = entire_Asso2 %>% group_by(Month) %>%
  summarise(UniqueWaterNames = n_distinct(waternames))

#sections
#Jan - 1
#Feb - 1
#March - 1
#April - 2
#May - 2
#June - 2
#July - 2
#August - 2
#Sept - 2
#Oct - 2
#Nov - 2
#Dec - 2


revisedcountsAsso2 = c(6,6,6,4,4,3,5,5,5,2,2,2)
RevisedSummaryAsso2 = cbind(Asso22022_summary,revisedcountsAsso2)
colnames(RevisedSummaryAsso2) = c("Month","Village", "Count", "RevisedCount")


```

Mailao 
```{r}
unique_sections_per_month_Mailao <- Mailao2022 %>%
  group_by(Month) %>%
  summarise(unique_sections = unique(sectionname))

entire_Mailao <- Mailao2022 %>%
  filter(treatmenttype == 'entire') %>%
  group_by(Month) %>%
  summarise(waternames = unique(watername))

#counts for entire 
entire_Mailaocounts = entire_Mailao %>% group_by(Month) %>%
  summarise(UniqueWaterNames = n_distinct(waternames))

#sections
#Jan - 1
#Feb - 1
#March - 1
#April - 1
#Sept - 1
#Nov - 1

revisedcountsMailao = c(8,7,7,5,17,22,18,8,13,9,3,6)
RevisedSummaryMailao = cbind(Mailao2022_summary,revisedcountsMailao)
colnames(RevisedSummaryMailao) = c("Month","Village", "Count", "RevisedCount")


```

Diganali
```{r}
unique_sections_per_month_Dig <- Diganali2022 %>%
  group_by(Month) %>%
  summarise(unique_sections = unique(sectionname))

#Diganali has no point of entry or sectioned water bodies 

entire_Dig <- Diganali2022 %>%
  filter(treatmenttype == 'entire') %>%
  group_by(Month) %>%
  summarise(waternames = unique(watername))

#counts for entire 
entire_Digcounts = entire_Dig %>% group_by(Month) %>%
  summarise(UniqueWaterNames = n_distinct(waternames))

revisedcountsDig = c(22,38,37,32,55,20,18,14,11,9,12) #numbers aren't exactly same as before, suggests that same water bodies were treated multiple times in the same month and being double counted before. this should be better now. 

RevisedSummaryDig = cbind(Diganali2022_summary,revisedcountsDig)
colnames(RevisedSummaryDig) = c("Month","Village", "Count", "RevisedCount")

```

Guizeme
```{r}
unique_sections_per_month_Gui <- Guimeze2022 %>%
  group_by(Month) %>%
  summarise(unique_sections = unique(sectionname))

entire_Gui <- Guimeze2022 %>%
  filter(treatmenttype == 'entire') %>%
  group_by(Month) %>%
  summarise(waternames = unique(watername))

#counts for entire 
entire_Guicounts = entire_Gui %>% group_by(Month) %>%
  summarise(UniqueWaterNames = n_distinct(waternames))

#there are no waterbodies treated in their entirety, only in sections or point of entry for this village

#Jan - 2
#Feb - 5
#March - 6
#April - 6
#May - 4
#June - 4
#July - 4
#August - 3
#Sept - 1
#Oct - 2
#Nov - 2
#Dec - 1

revisedcountsGui = c(2,5,6,6,4,4,4,3,1,2,2,1)
RevisedSummaryGui = cbind(Guimeze2022_summary,revisedcountsGui)
colnames(RevisedSummaryGui) = c("Month","Village", "Count", "RevisedCount")

#Dave agreed we should exclude this from the analysis as there are no waterbodies treated in their entirety and most are branches of the Chari 
```

Toukra Mousgam
```{r}
unique_sections_per_month_TM <- TM2022 %>%
  group_by(Month) %>%
  summarise(unique_sections = unique(sectionname))

entire_TM <- TM2022 %>%
  filter(treatmenttype == 'entire') %>%
  group_by(Month) %>%
  summarise(waternames = unique(watername))

#counts for entire 
entire_TMcounts = entire_TM %>% group_by(Month) %>%
  summarise(UniqueWaterNames = n_distinct(waternames))

#there are no waterbodies treated in their entirety, only in sections or point of entry for this village

#Jan - 1
#Feb - 1
#March - 1
#April - 1
#May - 2
#June - 2
#July - 1
#August - 1
#Sept - 1
#Oct - 1

revisedcountsTM = c(1,1,1,1,2,2,1,1,1,1)
RevisedSummaryTM = cbind(TM2022_summary,revisedcountsTM)
colnames(RevisedSummaryTM) = c("Month","Village", "Count", "RevisedCount")

```

Madjira 1
```{r}
unique_sections_per_month_Mad1 <- Madjira12022 %>%
  group_by(Month) %>%
  summarise(unique_sections = unique(sectionname))

entire_Mad1 <- Madjira12022 %>%
  filter(treatmenttype == 'entire') %>%
  group_by(Month) %>%
  summarise(waternames = unique(watername))

#counts for entire 
entire_Mad1counts = entire_Mad1 %>% group_by(Month) %>%
  summarise(UniqueWaterNames = n_distinct(waternames))

#there are no waterbodies treated in their entirety, only in sections or point of entry for this village

#Jan - 2
#Feb - 3
#March - 2
#April - 1
#May - 1
#June - 1
#July - 1
#August - 5
#Sept - 5
#Oct - 5
#Nov - 3
#Dec - 3

revisedcountsMAD = c(3,5,4,3,3,3,7,6,5,5,3,3)
RevisedSummaryMAD = cbind(Madjira12022_summary,revisedcountsMAD)
colnames(RevisedSummaryMAD) = c("Month","Village", "Count", "RevisedCount")

```

Kob
```{r}
unique_sections_per_month_kob <- kob2022 %>%
  group_by(Month) %>%
  summarise(unique_sections = unique(sectionname))

entire_kob <- kob2022 %>%
  filter(treatmenttype == 'entire') %>%
  group_by(Month) %>%
  summarise(waternames = unique(watername))

#counts for entire 
entire_kobcounts = entire_kob %>% group_by(Month) %>%
  summarise(UniqueWaterNames = n_distinct(waternames))

#there are no waterbodies treated in their entirety, only in sections or point of entry for this village

#April- 1
#June - 1
#Oct - 1
#Nov - 1
#Dec - 1

revisedcountskob = c(9,8,9,8,7,13,4,3,2,3,2,2)
RevisedSummarykob = cbind(kob2022_summary,revisedcountskob)
colnames(RevisedSummarykob) = c("Month","Village", "Count", "RevisedCount")

```
Plot data for each Village
```{r}
p1 <- ggplot(data = RevisedSummaryKS, aes(x = Month, y = RevisedCount)) +
  geom_point() + geom_line()

p2 <- ggplot(data = RevisedSummaryLC, aes(x = Month, y = RevisedCount)) +
  geom_point() + geom_line()

p3 <- ggplot(data = RevisedSummaryBB2, aes(x = Month, y = RevisedCount)) +
  geom_point() + geom_line()

p4 <- ggplot(data = RevisedSummaryAsso2, aes(x = Month, y = RevisedCount)) +
  geom_point() + geom_line()

p5 <- ggplot(data = RevisedSummaryMailao, aes(x = Month, y = RevisedCount)) +
  geom_point() + geom_line()

p6 <- ggplot(data = RevisedSummaryDig, aes(x = Month, y = RevisedCount)) +
  geom_point() + geom_line()

p7 <- ggplot(data = RevisedSummaryTM, aes(x = Month, y = RevisedCount)) +
  geom_point() + geom_line()

p8 <- ggplot(data = RevisedSummaryMAD, aes(x = Month, y = RevisedCount)) +
  geom_point() + geom_line()

p9 <- ggplot(data = RevisedSummarykob, aes(x = Month, y = RevisedCount)) +
  geom_point() + geom_line()

plot_grid(p1,p2,p3,p4,p5,p6,p7,p8,p9, labels = c('KS', 'LC','BB2','Asso2','Mailao','Dig','TM', 'Mad1','kob'))

ggplot(data = SelectVillagesRevised, aes(x = Month, y = RevisedCount,color = Village)) +
  geom_point() + geom_line()

```


Run Analysis for all villages: Asso 2, Mailao, Loumia Centre, Kolemara Sara, Bem-bem 2, Diganali, Guizeme (removed), Toukra Mousgam, Madjira 1, Kob
```{r}
SelectVillagesRevised = bind_rows(RevisedSummaryKS,RevisedSummaryLC,RevisedSummaryBB2,RevisedSummaryAsso2,RevisedSummaryMailao,RevisedSummaryDig,RevisedSummaryTM,RevisedSummaryMAD,RevisedSummarykob) 

SelectVillagesRevised$Village = as.factor(SelectVillagesRevised$Village)

SelectResults2022Revised = gamm(round(Count) ~ s(Month) + s(Village, bs = "re"),
                         family = quasipoisson(), correlation = corCAR1(form = ~Month | Village),
                         data = SelectVillagesRevised) 

SelectResults2022Revisedplot =plot_smooth(SelectResults2022Revised$gam, view="Month",lwd=2,
                                  transform=exp,se=1, shade=T, #ylim = c(0,50),
                                  ylab="Water Body Count",rug=F,
                                  hide.label=T, n.grid = 100)$fv

ggplot() + geom_line(data = SelectResults2022Revisedplot,aes(x=Month, y=fit)) +  #fitted line
  geom_ribbon(data = SelectResults2022Revisedplot,aes(x=Month, y=fit, ymin = ll,ymax=ul),alpha=0.2) + geom_point(data=SelectVillagesRevised,aes(x=Month,y=Count)) #raw data

```
Hierarchical GAMM
- factor-smoother interaction or"fs"
-m=1 for the group-level smoothers, which means the marginal TPRS basis for this term will penalize the squared first derivative of the function
-m=2, penalize the second derivative 
```{r}
SelectVillagesRevised$Village = as.factor(SelectVillagesRevised$Village)


Test1 = gam(round(Count) ~ s(Month, by=Village) + s(Village,bs="re"), 
                         family = quasipoisson(), data = SelectVillagesRevised, method = "REML") 


summary(Test1)

plot(Test1)


plot_smooth(Test1,view="Month",plot_all = "Village",transform = exp) 


Test2 = gam(round(Count) ~ s(Month) + s(Month, by=Village) + s(Village,bs="re"), 
                         family = quasipoisson(), data = SelectVillagesRevised, method = "REML") 

summary(Test2)

plot(Test2)

plot_smooth(Test2,view="Month",plot_all = "Village",transform = exp) 

plot_smooth(Test2,view="Month",transform = exp) 




#Bayesian method to account for singularity.
#Singularity occurs when model has too many parameters or terms, and some of them are redundant or perfectly correlated...In the context of GAMMs, singularity can occur when there is a perfect relationship between some of the smooth terms, making it impossible to estimate the unique contribution of each term. This issue often arises when using complex smoothing structures, and it can manifest as an error during the fitting process.

library(brms)
brm_model = brm(round(Count) ~ s(Month) + s(Month, by=Village) + s(Village,bs="re"),
                family = negbinomial(), data = SelectVillagesRevised, chains = 4, iter = 2000, warmup = 1000, control = list(max_treedepth = 20))

conditional_effects_plot <- conditional_effects(brm_model)
plot(conditional_effects_plot, points = TRUE, ask = FALSE)

#result: The largest R-hat is 1.93, indicating chains have not mixed.

brm_model_more_iterations <- update(brm_model, iter = 4000) 


```


Just look at Mailao
-for every waterbody, I want the name and treatment status for that month
-only one row if a water body if it's been treated in sections



